---
title: "How to use the updated `processFCC` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(processFCC)
```


Install and load the package with the following code:

```{r eval=FALSE, warning = FALSE}
install.packages("devtools")
library(devtools)
install_github("kdmulligan/processFCC", build_vignettes = TRUE)
library(processFCC)
```

One of the first things you will need to do is decide which time points of the 
FCC fixed broadband data you wish to use and process. Use `new_or_old_FCC()` to
check which dates are available and also the format (new or old) that should be 
used to process the data. The output of the function tells you the availables 
dates, the format, the function to use to process the data, and the website if
you would like to read more.

***explain why there are new and old formats***

```{r}
new_or_old_FCC()
```

Dates from December 2021 and prior use the old format whereas dates June 2022 
and more recent use the new format.

If you wish to use a time point of data in the "old" format you will need to use
the following functions: `old_FCC_links()`, `csv_to_sql_db()`, and 
`rollup_old_FCC()`.

For a time point of data in the "new" format you will need functions: 
`rollup_new_FCC()`. The simplest way to access the new data is to use the Public 
Data API. To do this you will have to create an account and setup an API key. 
Two of the arguments in `rollup_new_FCC()` are your FCC account username and
your API key so this step is necessary in order to use the function. To create 
an account go to https://broadbandmap.fcc.gov/login and then click "Create an 
account" in the bottom right of the sign-in box.


use this to right the instructions.

https://us-fcc.app.box.com/v/bdc-public-data-api-spec




