---
title: "How to use the updated `processFCC` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(processFCC)
```

## Loading the package

Install and load the package with the following code:

```{r eval=FALSE, warning = FALSE}
install.packages("devtools")
library(devtools)
install_github("kdmulligan/processFCC")
library(processFCC)
```

## Explanation of FCC Form History

The Federal Communications Commission (FCC) has collected data from all 
facilities-based broadband providers twice a year using Form 477. This form 
includes information about where and what type of Internet access is provided 
at the census block level.  However, this form was discontinued in in December 
2021. 

Instead, broadband providers now submit information about their broadband 
internet access services in the broadband data collection system. Within this 
system, data is reported more granularity than the census block level. For more 
information about the filing system, please reference the form resources [website](https://www.fcc.gov/economics-analytics/industry-analysis-division/form-477-resources).

Because of these changes in reporting, one of the first things you will need to 
do is decide which time points of the FCC fixed broadband data you wish to use 
and process. Dates from December 2021 and prior use the “old” format whereas 
dates June 2022 and beyond use the “new” format. For both formats, fixed 
broadband data is available at two time points per year: June and December. Use 
`new_or_old_FCC()` to check which dates are available and also the format (new or 
old) that should be used to process the data. The output of the function 
provides the available dates, the format, the function to use to process the 
data, and the website if you would like more information.

## New FCC data format

To process a data time point in the “new” format you will need functions: 
`avail_new_dates()` and `rollup_new_FCC()`. The most efficient way to work with 
the new FCC data is the public data application programming interface (API). 
Using the API functionality avoids the laborious task of downloading each data 
file individually from the website. 

To use the FCC API and the functions in this package for the “new” data format, 
it is necessary need to create an FCC User Registration account and then setup 
an API key. To create an account, go to <https://broadbandmap.fcc.gov/login> and 
then click “Create an account” in the bottom right of the sign-in box. 

Once logged into your account, click on your username in the top right corner. 
Then select the “Manage API Access” option. From this new page, click the 
“Generate” button to generate a new API token. Copy the token value and save it 
in a separate location. You will need this key and your username in 
`avail_new_dates()` and `rollup_new_FCC()` to access the database. For more 
information on creating an API key visit the 
[FCC API Instructions](https://us-fcc.app.box.com/v/bdc-public-data-api-spec).

Once your account and API key is set up.  The function `avail_new_dates()` can 
be used to see available dates in the new FCC data format.

```{r, eval=FALSE, warning = FALSE}
avail_new_dates(
  fcc_username = "email@location.com",
  api_key = "longstringofapicharacters"
)
```


If you already know what date you would like to process you can move right to 
using `rollup_new_FCC()`. The rollup functions count the number of unique 
broadband providers providing internet at or above the speed threshold 
combinations, up to 5 speed thresholds combinations can be specified. It is also 
possible to exclude specific broadband technologies or process only some states.

The arguments of `rollup_new_FCC()` are as follows:

- `fcc_username`: username for existing FCC account
- `api_key`: user’s unique API key for accessing FCC data. Generated within FCC account.
- `get_year`: the year of the FCC data to process.
- `get_month`: the month of the FCC data to process: either “Jun” or “Dec”.
- `states`: A vector of the state(s) abbreviations to include in the final data. 
The default, NULL, includes all states and territories in the final data set.
- `geogr`: Character representation of Census geography to summarize the data 
set at: census block (cb), census block group (cbg), census tract (ct), county 
(county).
- `tech_exc`: Vector of technology codes to exclude from data when rolling up. 
If you do not wish to exclude any technology codes input NA or c(NA). By default, 
satellite technologies are excluded.
- `thresh_down`: Vector of download speeds thresholds with maximum length of 5.
The vector must be the same length as `thresh_up` because elements of the vectors 
will be matched to count the number of internet providers at the given 
download/upload speed combinations.
- `thresh_up`: Vector of download speeds thresholds with maximum length of 5. 
The vector must be the same length as `thresh_down` because elements of the 
vectors will be matched to count the number of internet providers at the given 
download/upload speed combinations.
- `save_csv`: Logical for whether or not to save the processed data as a CSV.
- `wd`: filepath representing the working directory where the CSV should be saved. 
By default, this argument is set to the current working directory which is the 
file location in a qmd/rmd document or R project.

In the following call to `rollup_new_FCC()`, the June 2022 FCC data is rolled up 
to the census block level for all states (`states = NULL`) excluding technology codes 60, and 70. 
For each census block, the processed data considers 5 threshold combinations for 
download/upload speeds: 25/3, 25/5, 50/5, 75/10, and 100/100 Mbps. The processed 
data will count the number of providers within the census block providing 
broadband at the threshold speeds and excluding the specified technologies. No 
CSV file is saved with the final dataset.

```{r, eval=FALSE, warning = FALSE}
rollup_new_FCC(
  fcc_username = "email@location.com",
  api_key = "longstringofapicharacters",
  get_year = "2022",
  get_month = "Jun",
  states = NULL,
  geogr = "cb",
  tech_exc = c("60", "70"),
  thresh_down = c(25, 25, 50, 100, 100),
  thresh_up = c(3, 5, 10, 10, 100),
  save_csv = FALSE,
  wd = getwd()
)
```


## Old FCC data format

If you wish to use a time point of FCC data from the "old" format you will need to use
the following functions: `old_FCC_links()`, `csv_to_sql_db()`, and 
`rollup_old_FCC()`.

`old_FCC_links()` returns URL(s)for the download website for the unprocessed 
data set of interest. If year and month are left NULL then the output data set 
will contain all available years and months for the old format. Set most_recent
to TRUE in order to only get the link for the most recent version of the 
dataset(s), otherwise all available versions will be output. For example, June 
2018 is one time point of the data and there may be multiple versions of one 
time point as the data is updated.

The following code results in a data set with the URLs for where to download
the most recent versions of the June and December 2020 FCC data.

```{r, eval=FALSE, warning = FALSE}
old_FCC_links(year = 2020, month = NULL, most_recent = TRUE)
```

Based on the FCC data time point you would like to process, copy the
link from the `old_FCC_links()` output, go to the website, and download the US - 
Fixed with Satellite data set under the "Fixed Broadband Deployment Block Data" 
header. The code should still work if you decide to work with a state-level data 
set but the code is originally designed to work with the U.S. data set.
Once on the website from `old_FCC_links()`, the link for the csv file 
may take you to dropbox. In this case you will need to click on the 3 dots icon 
on the top bar of the page and then click download.

Next, the FCC CSV file should be added to a SQLite database. To do this, first, create the 
database connection with `dbConnect()`. Here, we name the database fcc.sqlite. You 
can use `dbListTables()` to check if any tables already exist in the database if 
you have used the connection previously. Next, in the call to `csv_to_sql_db()`, we add the FCC 
data for June 2020 in Alaska to the SQLite database. The name of the table being 
created in the SQLite database is fcc_ak_2020_Jun, the default table name would 
not include the Alaska designation. In `csv_to_sql_db()`, you must set the 
arguments `csv_file` and `con`.

```{r, eval=FALSE, warning = FALSE}
fcc_con <- dbConnect(SQLite(), dbname = "fcc.sqlite")
dbListTables(fcc_con)

csv_to_sql_db(csv_file = "C:/Users/kaile/Downloads/AK-Fixed-Jun2020-v2.csv", 
              con = fcc_con, 
              new_tbl_name = "fcc_ak_2020_Jun",
              year = 2020, month = "Jun",
              pre_process_size = 1000, chunk_size = 50000, 
              show_progress_bar = TRUE)
```

Once the data is in the SQLite database, we can then process it using 
`rollup_old_FCC()`. The output of this function is a data set based on the 
specifications from the function arguments similar to `rollup_new_FCC()`.
The following call to `rollup_old_FCC()` processes the June 2020 Alaska FCC data
that was just loaded to the SQLite database. The processed data is rolled up to 
the Census Tract level excluding technology codes 0, 60, and 70 and looks at 5 
threshold combinations for download/upload speeds: 25/3, 25/5, 50/5, 75/10, 
and 100/100 Mbps. The processed data will count the number of providers within 
the census tract providing broadband at the thresholds. No CSV file is saved 
with the final dataset.

In the `rollup_old_FCC()` function, the arguments `con` and `table_in_con` are 
the only two differences from `rollup_new_FCC()`: `con` is a SQLite database 
connection and `table_in_con` is the table in the `con` database to process.

```{r, eval=FALSE, warning = FALSE}
processed_dat <- rollup_old_FCC(
    con = fcc_con,
    table_in_con = "fcc_ak_2020_Jun",
    year = 2020, month = "Jun",
    geogr = "ct",
    tech_exc = c("0", "60", "70"),
    thresh_down = c(25, 25, 50, 75, 100),
    thresh_up = c(3, 5, 5, 10, 100)
  )
head(processed_dat)
```

If you are done working with the SQLite database, be sure to disconnect from the 
database with `dbDisconnect()`. You may also want to remove the sqlite, ZIP, 
and raw CSV files from your working directory because they are considerably 
large and take up a lot of space.

```{r, eval = FALSE}
dbDisconnect(con)
file.remove("fcc.sqlite")
```


